Enhanced Material Point Method to face dynamic problems: Local-maximum entropy approximation and explicit predictor-corrector scheme

### 2020-04-22 _region_.tex
Arreglando liadas

\message{ !name(main.tex)}%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments

gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
% flush right qed marks, e.g. at end of proof
\smartqed

\usepackage[lofdepth,lotdepth]{subfig}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{subfigure} % subfiguras
%
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{epstopdf}
\usepackage{mathrsfs} % para formato de letra
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{color}
\usepackage{makeidx}
\usepackage{xstring}
\usepackage{setspace}
\usepackage{epsfig}

\usepackage{pgf}
% \usepgfplotslibrary{external} 
% \tikzexternalize

%\usepackage{subfigure}
\usepackage{preview}
% Packages to write pseudo-algorithms %
\usepackage{algorithm}
\usepackage{algorithmic}

@@

%\usepackage{latexsym}
% etc.

%

% Matrix and vector nodes
\newcommand{\Matrix}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Vector}[1]{
  \ensuremath{\mathbf{{#1}}}
}

% Divergence
\newcommand{\Div}[1]{
  \ensuremath{div({#1})}
}
% Gradient
\newcommand\Grad[1]{grad({#1})}
\newcommand\GradS[1]{grad^s({#1})}
\newcommand\GradT[1]{grad^T({#1})}

% Partial derivative
\newcommand{\Deriv}[3][]{
  \ensuremath{\frac{\partial^{#1}{#2}}{ \partial {#3}^{#1} }}
}

% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\varGamma_d}{#2}\ d\varGamma}}
    {3}{\ensuremath{\int_{\varOmega}{#2}\ d\varOmega}}
  }
}

%
% Insert the name of "your journal" with
\journalname{Computational Particle Mechanics}
%
\begin{document}

\message{ !name(main.tex) !offset(-3) }

% \title{Novel improvements in the Material Point Method to face
%   dynamic problems: 
%   Local-maximum entropy approximation and an explicit predictor-corrector scheme. \thanks{Funding: This
%     study was funded by Agustín de Betancourt Foundation (grant number
%     262390106114).}
% }

\title{Enhanced Material Point Method to face
  dynamic problems:\thanks{Funding: This
    study was funded by Agustín de Betancourt Foundation (grant number
    262390106114).}
}
\subtitle{Local-maximum entropy approximation and explicit
  predictor-corrector scheme}

\titlerunning{Novel improvements in the Material Point Method to face
  dynamic problems.} % if too long for running head

\author{Miguel Molinos \and
  Pedro Navas \and
  Manuel Pastor \and
  Miguel Martín Stickle
}

%\authorrunning{Short form of author list} % if too long for running head

% \institute{Miguel Molinos \at
%   E.T.S de Ingenieros de Caminos, Canales y Puertos. \\
%   Universidad Politéctnica de Madrid, 28040 Madrid, Spain \\
%   \email{m.molinos@alumnos.upm.es}
%   % \emph{Present address:} of F. Author  %  if needed
%   \and
%   Pedro Navas 
%   \and
%   Manuel Pastor
%   \and
%   Miguel Martín Stickle
% }
% \institute{Miguel Molinos, Pedro Navas and Manuel Pastor \at E.T.S
%   de Ingenieros de Caminos, Canales y Puertos, Universidad Politéctnica de Madrid, 28040 Madrid, Spain, \; \email{m.molinos@outlook.es, pedro.navas@upm.es, manuel.pastor@upm.es}
% }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle
\begin{abstract}
  Material Point Method (MPM) has arisen in the recent years as an
  alternative to Finite Element Method (FEM) under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard MPM time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper, an explicit predictor-corrector time integration
  scheme has been proposed. Its superior performance mitigates the
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in MPM 
  occurs when to material points cross computational grid
  boundaries and motivated due to the lack of smoothness of the interpolation
  functions. This noise results in spurious local variations at the material
  points, where strain-stress fields are computed. This could invalidate
  the solution in one case, or damage it in another. To overcome it,
  this document adopts the \textit{local maximum-entropy approximation
    schemes} (LME) a robust substitute for the wide range of shape
  function in MPM. Local \textit{max-ent} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter they
  are able to behave finite element like or as a mesh-free
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
  \keywords{LME \and MPM \and Explicit predictor-corrector \and Dynamic problems}
  % \PACS{PACS code1 \and PACS code2 \and more}
  % \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

However, this method suffers other kind of instabilities, such those
when material points crossing cell boundaries. This give rise to the
development of other interpolation techniques to overcome this
limitation such as the generalized interpolation material point method
(GIMP) Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrate to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is  a voxel based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the dual domain material point method
(DDMP) proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}, therefore DDMP requires a large number
of particles needed for convergence, this
makes the method very expensive. In recent years the employ of
spline-lines has gain popularity with the introduction of the B-Spline
MPM proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employ of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to MPM framework with the Conservative Taylor
Least Squares (CTLS) reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}, unfortunately when particles are spread in
a challenging way, the quality of the CTLS approximation decrease
locally.

@@

\subsection{The boundary-value problem}
\label{sec:The boundary-value-problem}
In MPM the continuum mechanics approach is considered. So on, let
define a continuum domain $\varOmega$ occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and $\partial
\varOmega$ the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field $\vect{u}$ allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field $\vect{u}$ through the following set of governing equations and
restrictions that must be satisfied. First (i) the \textit{compatibility
  equation} that extracts from $\vect{u}$ the strain field is,
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type
$\varGamma_d$. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected. The corresponding
conjugate variable for the strain field if the stress field
$\tens{\sigma}$, which satisfies (ii) the \textit{conservation of
  momentum equation}
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type
$\varGamma_n$. Next (iii) the constitutive equation as a linear
application from $\Re^n$ to  $\Re^n$, which relates the
strain tensor with the stress tensor,
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

@@

Finally, to solve the equation \eqref{eq:particle_balance_forces3}, a second order
temporal integration scheme is required. Therefore, time is
discretized in to a finite set of time steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps. Once the nodal equilibrium equation it is solved, the values in the nodes are
interpolated back in to the particles and each particle it is advected
to the new position,
\begin{equation}
  \label{eq:Updated_Lagrangian}
  \dot{\vec{v}}_p = N_{Ip}\ \vec{a}_{I},\quad and\quad
  \dot{\vec{x}}_{p} = N_{Ip}\ \vec{v}_{I}  
\end{equation}
In MPM literature, this equations
(\ref{eq:particle_balance_forces3}) and (\ref{eq:Updated_Lagrangian}),
are solved with an explicit forward Euler algorithm. The time integration scheme of MPM has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009} and summarized
in figure~\ref{fig:MPM_algorithm}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme}
  }
  \caption{Description of MPM standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Other authors have proposed many others time integration alternatives
like \cite{Guilkey2003}, \cite{Tran2019e}, \cite{Charlton2017}. In the
first publication on MPM \cite{Sulsky1994}, the nodal acceleration
was employed to update the particles as
\begin{equation}
  \label{eq:Sulsky-1994-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\ N_{Ip}^{k}\ \vec{a}_{I}^{k}
\end{equation}
\begin{equation}
  \label{eq:Sulsky-1994-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ N_{Ip}^{k}\ \vec{v}_{I}^{k}.
\end{equation}
However, as Andersen (2009)\cite{thesis_Andersen_2009} point out, this algorithm has been shown to be numerically unstable due to that
$\vect{f}_I^{int,k}$ can be finite for an infinitesimal nodal mass
$\tens{m}$. This can lead to numerical issues when nodal acceleration
is obtained for evaluating \eqref{eq:Sulsky-1994-UL-x},\eqref{eq:Sulsky-1994-UL-v}. Hence, a
corrected version of this algorithm in shown in Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}
\begin{equation}
  \label{eq:Zhang-2016-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \frac{N_{Ip}^{k}\ \vec{p}_{I}^{k}}{\tens{m}_I}.  
\end{equation}
\begin{equation}
  \label{eq:Zhang-2016-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\ \frac{N_{Ip}^{k}\ \vec{f}_{I}^{k}}{\tens{m}_I},
\end{equation}
Later Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for MPM inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both in the beginning and the end of the
time step.
\begin{equation}
  \label{eq:Tran-2019-GA-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\  N_{Ip}^{k}\ \left[(1 - \gamma)\ \vect{a}_I^{k} +
    \gamma\ \vect{a}_I^{k+1} \right],\\
\end{equation}
\begin{equation}
\label{eq:Tran-2019-GA-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + N_{Ip}^{k} \left[ \Delta t\ \vec{v}_{I}^{k}+ \Delta t^2\left( (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right) \right]
\end{equation}
\begin{equation}
  \label{eq:Tran-2019-GA-a}
  \vect{a}_p^{k+1} = N_{Ip}^{k}\ \vec{a}_{I}^{k+1}.
\end{equation}

@@

\cite{Arroyo2006} (e.g. two dimensional example shown in figure~\ref{fig:LME_MPM}).
 Since  $p_I$  is defined in the entire domain, in practice, the

@@

neighbours, see \cite{Arroyo2006} for details. This tolerance defines
the limit values of the influence radius and is used thereafter to
find the neighbour nodes of a given integration point. Regarding this
aspect, an crucial consideration for the implementation of the LME
shape functions in MPM is exposed. To avoid the transference of
particle information to those nodes fare apart from the particle
domain $\Omega_p$, a especial treatment of the neighbour nodes for
those particles close to the boundary is required. It consists in
select only those nodes which position $\vect{x}_I$ lies inside of the
domain $\Omega$. This can be performed by employing the nodal
connectivity of the background mesh. And activating for computing only
those nodes whose element has a particle inside of it as we can see in
figure~\ref{fig:Particle-discretization}. 

@@

@@

      \includegraphics[width=0.16\textwidth]{Figures/MPM_Shape_Fun}\\
      \includegraphics[width=0.16\textwidth]{Figures/MPM_Shape_Fun_dx}\\
      \includegraphics[width=0.16\textwidth]{Figures/MPM_Shape_Fun_dy}

@@

      \includegraphics[width=0.16\textwidth]{Figures/LME_17_3_Shape_Fun}\\
      \includegraphics[width=0.16\textwidth]{Figures/LME_17_3_Shape_Fun_dx}\\

>>>>

\message{ !name(LME-MPM.tex)}% %%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
% %
% % This is a general template file for the LaTeX package SVJour
% % for Springer journals.          Springer Heidelberg 2010/09/16
% %
% % Copy it to a new file with a new name and use it as the basis
% % for your article. Delete % signs as needed.
% %
% % This template includes a few options for different layouts and
% % content for various journals. Please consult a previous issue of
% % your journal as needed.
% %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %
% % First comes an example EPS file -- just ignore it and
% % proceed on the \documentclass line
% % your LaTeX will extract the file if required
% \begin{filecontents*}{example.eps}
% %!PS-Adobe-3.0 EPSF-3.0
% %%BoundingBox: 19 19 221 221
% %%CreationDate: Mon Sep 29 1997
% %%Creator: programmed by hand (JK)
% %%EndComments

% gsave
% newpath
%   20 20 moveto
%   20 220 lineto
%   220 220 lineto
%   220 20 lineto
% closepath
% 2 setlinewidth
% gsave
%   .4 setgray fill
% grestore
% stroke
% grestore
% \end{filecontents*}
% %
% \RequirePackage{fix-cm}
% %
% %\documentclass{svjour3}                     % onecolumn (standard format)
% %\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
% %\documentclass[smallextended]{svjour3}       % onecolumn (second format)
% \documentclass[twocolumn]{svjour3}          % twocolumn
% %
% % flush right qed marks, e.g. at end of proof
% \smartqed

% \usepackage[lofdepth,lotdepth]{subfig}
% \usepackage{graphicx}
% \usepackage{subcaption}

% \usepackage{amsmath,amsthm,amssymb}

% \usepackage{amsfonts}
% \usepackage{epstopdf}
% \usepackage{mathrsfs} % para formato de letra
% \usepackage{hyperref}
% \usepackage{enumitem}
% \usepackage{array}
% \usepackage{tabularx}
% \usepackage{supertabular}
% \usepackage{fancyhdr}
% \usepackage{multirow}
% \usepackage{color}
% \usepackage{makeidx}
% \usepackage{setspace}
% \usepackage{epsfig}

% \usepackage{pgf}
% % \usepgfplotslibrary{external} 
% % \tikzexternalize

% %\usepackage{latexsym}

\documentclass[preprint,12pt,a4paper]{elsarticle}

\usepackage{lineno,hyperref}
\usepackage{float}
\usepackage{subfig}
\usepackage{color}
\usepackage{soul}

% Packages to write pseudo-algorithms %
\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage{cancel}

\usepackage{subfigure} % subfiguras

\usepackage{amsmath,amsthm,amssymb}
\usepackage{xstring}

@@

\newcommand{\vec}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\tens}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Matrix}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Vector}[1]{
  \ensuremath{\mathbf{{#1}}}
}
% Divergence
\newcommand{\Div}[1]{
  \ensuremath{div({#1})}
}
% Gradient
\newcommand\Grad[1]{grad({#1})}
\newcommand\GradS[1]{grad^s({#1})}
\newcommand\GradT[1]{grad^T({#1})}
% Partial derivative
\newcommand{\Deriv}[3][]{
  \ensuremath{\frac{\partial^{#1}{#2}}{ \partial {#3}^{#1} }}
}
% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\varGamma_d}{#2}\ d\varGamma}}
    {3}{\ensuremath{\int_{\varOmega}{#2}\ d\varOmega}}
  }
}

\modulolinenumbers[5]

\journalname{Computer Methods in Applied Mechanics and Engineering}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\message{ !name(LME-MPM.tex) !offset(-3) }

\begin{frontmatter}

\title{Enhanced Material Point Method to face
  dynamic problems: Local-maximum entropy approximation and explicit
  predictor-corrector scheme}

%% Group authors per affiliation:
\author{
Miguel Molinos$^a$
Pedro Navas$^a$\footnote{Corresponding author: p.navas@upm.es},
Manuel Pastor$^a$
and Miguel Martín Stickle$^a$ 
}
\address{
  $^a$ ETSI Caminos, Canales y Puertos, Universidad Polit\'ectnica de Madrid.\\ c. Prof. Aranguren 3, 28040 Madrid, Spain
}

\begin{abstract}
  Material Point Method (MPM) has arisen in the recent years as an
  alternative to Finite Element Method (FEM) under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard MPM time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper, an explicit predictor-corrector time integration scheme has been proposed. Its superior performance mitigates the 
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in MPM 
  occurs when to material points cross computational grid
  boundaries and motivated due to the lack of smoothness of the interpolation
  functions. This noise results in spurious local variations at the material
  points, where strain-stress fields are computed. This could invalidate
  the solution in one case, or damage it in another. To overcome it,
  this document adopts the \textit{local maximum-entropy approximation
    schemes} (LME) a robust substitute for the wide range of shape
  function in MPM. Local \textit{max-ent} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter they
  are able to behave finite element like or as a mesh-free
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
\end{abstract}

\begin{keyword}
  LME \sep MPM \sep Explicit predictor-corrector \sep Dynamic problems}
\end{keyword}

\end{frontmatter}

\linenumbers

@@

However, this method suffers other kind of instabilities, such those
when material points crossing cell boundaries. This give rise to the
development of other interpolation techniques to overcome this
limitation such as the generalized interpolation material point method
(GIMP) Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrate to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is  a voxel based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the dual domain material point method
(DDMP) proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}, therefore DDMP requires a large number
of particles needed for convergence, this
makes the method very expensive. In recent years the employ of
spline-lines has gain popularity with the introduction of the B-Spline
MPM proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employ of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to MPM framework with the Conservative Taylor
Least Squares (CTLS) reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}, unfortunately when particles are spread in
a challenging way, the quality of the CTLS approximation decrease
locally.

@@

\subsection{The boundary-value problem}
\label{sec:The boundary-value-problem}
In MPM the continuum mechanics approach is considered. So on, let
define a continuum domain $\varOmega$ occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and $\partial
\varOmega$ the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field $\vect{u}$ allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field $\vect{u}$ through the following set of governing equations and
restrictions that must be satisfied. First (i) the \textit{compatibility
  equation} that extracts from $\vect{u}$ the strain field is,
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type
$\varGamma_d$. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected. The corresponding
conjugate variable for the strain field if the stress field
$\tens{\sigma}$, which satisfies (ii) the \textit{conservation of
  momentum equation}
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type
$\varGamma_n$. Next (iii) the constitutive equation as a linear
application from $\Re^n$ to  $\Re^n$, which relates the
strain tensor with the stress tensor,
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

@@

Finally, to solve the equation \eqref{eq:particle_balance_forces3}, a second order
temporal integration scheme is required. Therefore, time is
discretized in to a finite set of time steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps. Once the nodal equilibrium equation it is solved, the values in the nodes are
interpolated back in to the particles and each particle it is advected
to the new position,
\begin{equation}
  \label{eq:Updated_Lagrangian}
  \dot{\vec{v}}_p = N_{Ip}\ \vec{a}_{I},\quad and\quad
  \dot{\vec{x}}_{p} = N_{Ip}\ \vec{v}_{I}  
\end{equation}
In MPM literature, this equations
(\ref{eq:particle_balance_forces3}) and (\ref{eq:Updated_Lagrangian}),
are solved with an explicit forward Euler algorithm. The time integration scheme of MPM has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009} and summarized
in figure~\ref{fig:MPM_algorithm}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme_horizontal}
  }
  \caption{Description of the three steps in MPM standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Other authors have proposed many others time integration alternatives
like \cite{Guilkey2003}, \cite{Tran2019e}, \cite{Charlton2017}. In the
first publication on MPM \cite{Sulsky1994}, the nodal acceleration
was employed to update the particles as
\begin{equation}
  \label{eq:Sulsky-1994-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\ N_{Ip}^{k}\ \vec{a}_{I}^{k}
\end{equation}
\begin{equation}
  \label{eq:Sulsky-1994-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ N_{Ip}^{k}\ \vec{v}_{I}^{k}.
\end{equation}
However, as Andersen (2009)\cite{thesis_Andersen_2009} point out, this algorithm has been shown to be numerically unstable due to that
$\vect{f}_I^{int,k}$ can be finite for an infinitesimal nodal mass
$\tens{m}$. This can lead to numerical issues when nodal acceleration
is obtained for evaluating \eqref{eq:Sulsky-1994-UL-x},\eqref{eq:Sulsky-1994-UL-v}. Hence, a
corrected version of this algorithm in shown in Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}
\begin{equation}
  \label{eq:Zhang-2016-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \frac{N_{Ip}^{k}\ \vec{p}_{I}^{k}}{\tens{m}_I}.  
\end{equation}
\begin{equation}
  \label{eq:Zhang-2016-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\ \frac{N_{Ip}^{k}\ \vec{f}_{I}^{k}}{\tens{m}_I},
\end{equation}
Later Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for MPM inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both in the beginning and the end of the
time step.
\begin{equation}
  \label{eq:Tran-2019-GA-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\  N_{Ip}^{k}\ \left[(1 - \gamma)\ \vect{a}_I^{k} +
    \gamma\ \vect{a}_I^{k+1} \right],\\
\end{equation}
\begin{equation}
\label{eq:Tran-2019-GA-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + N_{Ip}^{k} \left[ \Delta t\ \vec{v}_{I}^{k}+ \Delta t^2\left( (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right) \right]
\end{equation}
\begin{equation}
  \label{eq:Tran-2019-GA-a}
  \vect{a}_p^{k+1} = N_{Ip}^{k}\ \vec{a}_{I}^{k+1}.
\end{equation}

@@

\cite{Arroyo2006}, where $h$ is defined as a measure of the nodal
spacing. 
Since $N_I$ is defined in the entire domain, in practice, the

@@

neighbours, see \cite{Arroyo2006} for details.
This tolerance defines the limit values of the influence radius and is used thereafter to
find the neighbour nodes of a given integration point. An aditional remark is that in similariti  to alternative non-polynomial meshfree basis functions, the LME approximation scheme requires more than $d+1$ nodes to determine the values of the shape functions as well as their derivatives at any point in the convex hull of the nodal set, where $d$ is the dimension of the problem.

This interpolation technique avoid one important shortcoming when using
GIMP or B-Spline MPM regarding the computational domain boundaries
Steffen {\it et al.} (2008)\cite{Steffen2008b}. Which is the
additional considerations in the applications of the boundary
conditions. This is motivated because of their increased extents, it
is possible for particles to influence, and be influenced by, nodes
that lie outside of the simulation domain. Some researchers have solve
this problems with ``extra'', or the called ``ghost'' nodes by other investigators. Anyway this nodes requires especial treatments similar
to those employed in the smoothed particle hydrodynamics (SPH), for
further details see Liu \& Liu (2003)\cite{Liu2003}. The approach here
described does not requires the employ of this artifacts.

@@

  \resizebox{0.5\hsize}{!}{

@@

      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun_dy}

@@

      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun_dx}\\
